Hydras \& Co.: Formalized mathematics in Coq\\ for inspiration and entertainment

### 2021-11-23 doc/JFLA2022/paper.tex
Add reference to latest release in footnote.

We believe the \Hydras electronic book has found an attractive set of tradeoffs between the properties of executable books and traditional books. Specifically, in an executable book, the proof assistant language places limits on the structure. For example, Coq may not permit repeating a definition from another file verbatim, since the definition uses variables that are not present in the current context. Thanks to our \alectr based toolchain, the \Hydras book can include ``live'' code fragments at any point in the text. And in contrast to traditional books, we validate the consistency of code fragments during every build of the electronic book. On the one hand, we follow executable books in providing continuous delivery of new revisions at a high pace. But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.

The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the code, and our use of snippet delimiters in the Coq source files may lower source readability. For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce overhead without providing complementary benefits.
%
More specifically, we currently distribute the book as a PDF generated by LaTeX that links to corresponding HTML documentation of the Coq sources generated by \coqdoc. In contrast, executable books such as Software Foundations can provide a single uniform HTML representation of both book contents and code.

@@

We plan to extend \Hydras in two main directions. Firstly, we plan to bridge the combinatorial results of \Hydras and the set-theoretic content of \gaia, enabling the transfer of many interesting theorems between the two packages. Secondly, we aim to write a formal proof in \coq of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
%
Moreover, \Hydras is not limited to the study of ordinal numbers and applications.
We are also developing a package about efficient exponentiation
algorithms, and aim to eventually include new topics. We invite new collaborators to join us in our efforts.

>>>>

We believe the \Hydras electronic book has found an attractive set of tradeoffs between the properties of executable books and traditional books. Specifically, in an executable book, the proof assistant language places limits on the structure. For example, Coq may not permit repeating a definition from another file verbatim, since the definition uses variables that are not present in the current context. Thanks to our \alectr based toolchain, the \Hydras book can include ``live'' code fragments at any point in the text. And in contrast to traditional books, we validate the consistency of code fragments during every build of the electronic book. On the one hand, we follow executable books in providing continuous delivery of new revisions at a high pace. But on the other hand, we also aim to make regular timestamped versions compatible with specific versions of Coq, akin to new editions of traditional books.\footnote{The latest Hydras \& Co. release is 0.5, available at \url{https://github.com/coq-community/hydra-battles/releases/tag/v0.5}.}

The flip side of our approach is that we separate the Coq sources from the LaTeX sources that describe the code, and our use of snippet delimiters in the Coq source files may lower source readability. For books and libraries with small scope and specialized audiences, distributing knowledge representations between sources and documentation formats the way we do in \Hydras may introduce overhead without providing complementary benefits.
%
More specifically, we currently distribute the book as a PDF generated by LaTeX that links to corresponding HTML documentation of the Coq sources generated by \coqdoc. In contrast, executable books such as Software Foundations can provide a single uniform HTML representation of both book contents and code. We hope that future toolchain improvements can address all these issues.

@@

We plan to extend \Hydras in two main directions. Firstly, we plan to bridge the combinatorial results of \Hydras and the set-theoretic content of \gaia, enabling the transfer of many interesting theorems between the two packages. Secondly, we aim to write a formal proof in \coq of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
%
Moreover, \Hydras is not limited to the study of ordinal numbers and their applications.
We are also developing a package about efficient exponentiation
algorithms, and aim to eventually include new topics. We invite new collaborators to join us in our efforts.
